### 2019-10-15 Writing/Palaeontology/Sup1-CharDiff.tex
Prepared Palaeontology submission

@@

\section{Distance metric proof}

The Character Difference ($CD$) metric is a distance $d: CD \times CD \rightarrow [0;\infty)$ with any character $x$, $y$ and $z$ (using the \textit{xyz} notation) $\in CD$ if an only if the followings are satisfied:

\begin{equation}
    \label{Positivity}
    CD_{(x,y)} \geq 0
\end{equation}
\begin{equation}
    \label{Identity}
    CD_{(x,y)} = 0 \Leftrightarrow x = y
\end{equation}
\begin{equation}
    \label{Equality}
    CD_{x,y} = CD_{y,x}
\end{equation}
\begin{equation}
    \label{Subadditivity}
    CD_{(x,y)} \leq CD_{(x,z)} + CD_{(z,y)}
\end{equation}

\noindent Furthermore, $CD$ is a evolutionary distance metric if it satisfies the four point condition:

\begin{equation}
    \label{fourpoint}
    CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )}
\end{equation}

\newpage

\subsection{Character difference positivity}

For any $n \in [2;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$ because the difference between character states is considered as Fitch-like (unordered; i.e. the difference between two different characters states is equal to $1$ and the difference between two identical character states is equal to $0$).
Furthermore, following the $xyz$ notation described above, for any $n > 1$, $x_{1}-y_{1}=0$.
Therefore the sum of the differences $\sum_{i=1}^{n}\abs{x_{i}-y_{i}} \in [0;(n-1)]$.

\begin{equation}
    \label{positive_demo}
    \begin{gathered}
    0 \leq  \abs{x_{n}-y_{n}} \leq 1\\
    0 \leq  \sum_{i=1}^{n}\abs{x_{i}-y_{i}} \leq (n-1)\\
    0 \leq  \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} \leq 1\\
    \end{gathered}
\end{equation}

Thus equation \ref{Positivity} is true.

\newpage

\subsection{Character difference identity}

For any $n \in [2;\infty)$ and $x_{n} = y_{n}$ then:
%TG: Is that valid for a demonstration?
\begin{equation}
    \label{identitydemo}
    \begin{aligned}
    \abs{x_{n}-y_{n}} &= 0\\
    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= 0\\
    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} &= 0\\
    \end{aligned}
\end{equation}

Thus equation \ref{Identity} is true.

\subsection{Character difference equality}

For any $n \in [2;\infty)$ and $x_{n}$ and $y_{n} \in (0,1)$, similarly to equation \ref{positive_demo}:

\begin{equation}
    \label{equality_demo}
    \begin{aligned}
    \abs{x_{n}-y_{n}} &= \abs{y_{n}-x_{n}}\\
    \sum_{i}^{n}\abs{x_{i}-y_{i}} &= \sum_{i}^{n}\abs{y_{i}-x_{i}}\\
    \frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1} &= \frac{\sum_{i}^{n}\abs{y_{i}-x_{i}}}{n-1}\\
    \end{aligned}
\end{equation}

Thus equation \ref{Equality} is true.

\subsection{Character difference subadditivity}

For any $n, \in [2;\infty)$ and $x_{n}$, $y_{n}$, $z_{n}$ $\in (0,1)$:

\begin{equation}
    \label{equality_demo}
    \begin{aligned}
    \text{since:}\\
    \underbrace{\frac{\sum_{i}^{n}\abs{x_{i}-y_{i}}}{n-1}}_{\in[0,1]}
    &\leq
    \underbrace{\frac{\sum_{i}^{n}\abs{x_{i}-z_{i}}}{n-1}}_{\in[0,1]}
    +
    \underbrace{\frac{\sum_{i}^{n}\abs{z_{i}-y_{i}}}{n-1}}_{\in[0,1]}\\
    \text{thus:}\\
    \in[0,1] &\leq \in[0,1] + \in[0,1] \\
    \in[0,1] &\leq \in[0,2]\\
    \end{aligned}
\end{equation}
Thus equation \ref{Subadditivity} is true.

\newpage

\subsection{Four point condition}

For any $n \in [2;\infty)$ and any four distinct characters $i$, $j$, $n$ and $m$, and if equations \ref{Positivity} and \ref{Subadditivity} are true, the pairwise distances $CD_{(i,j)}$, $CD_{(m,n)}$, $CD_{(i,m)}$, $CD_{(j,n)}$, $CD_{(j,m)}$ and $CD_{(i,n)}$ can be represented as a combination of any five values $a$, $b$, $c$, $d$ and $e$ $\geq 0$:

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.5\textwidth]{Four_point.pdf}
\end{figure}

Then:

\begin{equation}
    \begin{aligned}
    CD_{(i,j)} = a + b \\
    CD_{(m,n)} = d + e \\
    CD_{(i,m)} = a + c + d \\
    CD_{(j,n)} = b + c + e \\
    CD_{(j,m)} = b + c + d \\
    CD_{(i,n)} = a + c + e \\
    \end{aligned}
\end{equation}

Thus:

\begin{equation}
    \begin{gathered}
    CD_{(i,j)} + CD_{(m,n)} \leq \text{max}{(CD_{(i,m)} + CD_{(j,n)} ; CD_{(j,m)} + CD_{(i,n)} )} \\ 
    a + b + d + e \leq \text{max} (a+c+d+b+c+e; b+c+d+a+c+e) \\
    a + b + d + e \leq a+b+2c+e+d \\
    2c \geq 0 \\
    \end{gathered}
\end{equation}

Thus equation \ref{fourpoint} is true.

>>>>

\noindent{\Large \bf Supplementary material 1 - Character difference}

@@

### 2019-10-15 Writing/Palaeontology/Tab_pooledsmethods_test.tex
Prepared Palaeontology submission

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Fri May 31 14:42:25 2019
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100 vs. c350 & \textbf{0.965} & 48263.500 & \textbf{0.000} \\ 
   & c100 vs. c1000 & 0.914 & 37714.500 & \textbf{0.000} \\ 
   & c350 vs. c1000 & \textbf{0.976} & 53138.000 & \textbf{0.000} \\ 
  Tr & c100 vs. c350 & 0.929 & 40253.000 & \textbf{0.000} \\ 
   & c100 vs. c1000 & 0.826 & 27923.000 & \textbf{0.000} \\ 
   & c350 vs. c1000 & \textbf{0.957} & 48164.500 & \textbf{0.000} \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the three different number of characters: 100 (c100), 350 (c350) and 1000 (c1000). Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscharacters_test}
\end{table}

>>>>

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Fri May 31 14:42:25 2019
\begin{table}[ht]
\centering
\begin{tabular}{llr|rr}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & bayesian vs. parsimony & 0.939 & 192659.000 & \textbf{0.000} \\ 
  Tr & bayesian vs. parsimony & \textbf{0.986} & 157677.500 & \textbf{0.041} \\ 
   \hline
\end{tabular}
\caption{Differences between the Normalised Tree Similarities in terms of clade conservation (Robinson-Foulds distances - RF) and taxa displacement (Triplets - Tr) for the two different inference methods: bayesian or maximum parsimony. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric Wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledsmethods_test}
\end{table}

### 2019-10-15 Writing/Palaeontology/draft.tex
Prepared Palaeontology submission

\documentclass[12pt,letterpaper]{article}
\usepackage{natbib}

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{outline.eps}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using MP and BPP methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines). $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
\label{Fig:outline}
\end{figure}

>>>>

% https://www.palass.org/sites/default/files/media/publications/for_authors/ITA_2016_v1.pdf
% Abstract < 300w ---> OK
% Keywirds < 6 ---> OK
\documentclass[12pt,letterpaper]{article}
\usepackage{natbib}

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{outline.eps}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using maximum parsimony and Bayesian inference methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines). $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
\label{Fig:outline}
\end{figure}
